### 2022-06-13 paper.tex

>>>>

\subsection{Simon's draft}

Here, we have proposed an improved strategy to preprocess single-cell bisulfite-sequencing data. On the one hand, we suggest a scoring scheme that is aware of a reads local context rather than just treating all reads in an interval alike, and on the other hand, we show a way to pinpoint minimal regions of high variability across cells, which we call variably-methylated regions (VMRs). By reanalyzing two published datasets we could show that these improvements to data preprocessing help to increase signal and descrease noise, resulting in a more informative intermediate-dimensional representation of the data. As examples of practical benefits, we demonstrate that our preprocessing allows for better distinction of cell sub-types, especially for smaller data sets with limited numbers of cells. In addition to this improvement of global signal-to-noise ratio (with regards to cell dissimilarity quantification), we also show that the determination of VMRs improve biological interpretability by pinpointing the regions that actually change and hence have a putative regulatory role.

Furthermore, we have presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms. The package also offers functionality for all other steps of data preprocessing taht are needed to get from the output of methylation callers such as Bismark and methylpy to a data matrix that is suitable as input for standard analysis tools that work in a reduced-dimensionality setting, thus allowing the user to use, e.g., well established tools from sc-RNA-Seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses etc. By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis. In addition, by pinpointing exact regions of change, the VMR functionality provides useful input for genomics-style analyses.

In conclusion, we have presented powerful improvements to sc-BS-Seq data preprocessing and a software tool that implements these and also completes a gap in existing tool chains.

\subsection{Lukas's draft}
